TINET+TECS: Component-based TCP/IP Protocol Stack for Embedded Systems

### 2017-02-07 ISORC2017/diff/20170207.tex
Add the current version to diff

Embedded systems are applied to Internet of Things (IoT), and the high productivity of embedded network software is required.
TINET is a TCP/IP protocol stack for embedded systems.
Although TINET is a compact TCP/IP protocol stack, it consists of many complex source codes.
Therefore, it is difficult to maintain, extend, and analysis the software.
To improve the scalability and configurability, this paper has proposed TINET+TECS, a component-based TCP/IP protocol stack for embedded systems: TINET componentized with TOPPERS embedded component system (TECS).
The component-based TINET provides software developers high productivity such as change of network buffer size and adding/removing TCP (or UDP) function.
We evaluate the component-based TINET compared with the original TINET.
We confirm that the overheads of execution time and memory consumption are low, and that the configurability is improved.

@@

This paper proposes component-based TCP/IP protocol stack, i.e., TINET+TECS, to improve the configurability and scalability of software.
TECS (TOPPERS Embedded Component System) \cite{par:TECS} is utilized to componentize TINET, because TECS is a component system suitable for embedded systems.
TECS supports static configuration, which statically defines component behaviors and interconnections, thus TECS can optimize the overhead of componentization.

In addition, the proposed framework utilizes dynamic connection, a novel method of TECS, to switch component bindings dynamically.
A general TCP/IP protocol server dynamically processes the requested port i.e., HTTP (port: 80), HTTPS (prot: 443).
However, embedded systems can not dynamically generate components due to the strict memory restriction.
TINET+TECS statically generates components, and dynamically combines them.
Therefore,

In the proposed framework, software application can be developed by not only TECS method but also existing method.
Software appication can be developed as a TECS component because TINET+TECS is a component-based framework using TECS.
Moreover, TECS supports an adapter to call functions of TECS components from non TECS codes.
The legacy codes such as an existing application can be supported in the proposed framework.

@@

This section describes system model, including the basic technologies such as TINET and TECS.
System model of the proposed framework is shown as Fig. \ref{fig:SystemModel}.
TINET+TECS is a component-based TCP/IP protocol stack, so TCP output task (tTCPOutput) and Ethernet input task (tEhternetInput) are implemented as TECS components.
CEP and REP (Section \ref{sec:TINET}), which are also implemented as TECS components, dynamically switch bindings by TECS method.
Moreover, TECS adapter supports the legacy codes such as an existing application.

\begin{figure}[t]
    \centering
    \includegraphics[width=7.0cm,clip]{figure/SystemModel.pdf}
    \caption{System model}
    \label{fig:SystemModel}
\end{figure}

@@

The components of protocol stack for TINET+TECS is shown as Fig. \ref{fig:ComponentProtocolStack}.
Note that some small particle components such as a kernel object, dataqueques, and semaphores are omitted to simplify the component diagram.
In TINET+TECS, the components are divided for each protocol, and the functionalities such as input function and output function are defined as each a component.
Therefore, software visibility is improving because of small grain components.

@@

{\bf Application layer:}
An application in TINET+TECS is implemented as a component such as tApplication.
Software with TINET uses ITRON TCP/IP API such as {\it tcp\_snd\_buf} and {\it tcp\_rcv\_buf}.
In TINET+TECS, the application component calls TECS functions such as {\it cTCPAPI\_sendBuffer} and {\it cTCPAPI\_receiveBuffer}.
Moreover, TINET+TECS supporting a TECS adapter (\ref{sec:TECS Adapter}), an exiting application with TINET can run on TINET+TECS framework without transporting.
Therefore, software can be developed either with exiting methods or as a TECS component.

{\bf Transport layer:}
tTCPCEP (tUDPCEP) is a CEP compornent for TCP (UDP), and tREP is a REP component.
For example, a server program supporting multiple clients can be developed by preparing the multiple tTCPCEP components.
tTCPInput (tUDPInput) and tTCPOutput (tUDPOutput) are components performing the receiving and sending processing respectively in the transport layer.

@@

To utilize the protocol stack as same as the original TINET, communication object components such as tTCPCEP, tUDPCEP, and tREP are defined as an interface between TINET+TECS and an application.
The communication object component is a component corresponding to a CEP or REP of the original TINET.
An application developers can utilize the TINET+TECS functionalities by generating and combining as many components as necessary.

@@

The original TINET eliminates dynamic memory control to meet the sever memory restriction of embedded systems.
A memory area for sending/receiving data in the protocol stack is allocated and released within an predetermined area.
Memory allocator component performs the elimination of dynamic memory control in TINET+TECS.
The component provides a requested memory area from the memory area statically allocated.

@@

{\it send} is also a specifier for transferring data to a callee from a caller such as {\it in}.
The difference between {\it in} and {\it send} is whether to deallocate the data memory in a caller or callee, shown as Fig. \ref{fig:SendReceive}.
In case of {\it in} specifier, both allocating and deallocating the data memory are performed in the caller.
On the other hand, in case of {\it send}, the caller allocates the data memory and the callee deallocates it.

@@

{\it receive} is also a specifier for a caller receiving data from a callee such as {\it out}.
The difference between {\it out} and {\it receive} is whether to allocate the data memory in a caller or callee, shown as Fig. \ref{fig:SendReceive}.
While, in case of {\it out}, a callee writes data in the memory allocated by a caller, in case of {\it receive}, the callee allocates the data memory.
Deallocating the memory is performed in the caller in both cases.

@@

\subsubsection{Network timer component}

@@

TECS supports dynamic connection method as a novel functionality.
Dynamic connection is a method to switch the binding of components at runtime as shown in Fig. \ref{fig:DynamicConnection}.
Note that all components are statically generated in TECS.
TECS can optimize the overhead of componentization because components are statically configured.
Dynamically generating the components causes a lot of memory consumption, which is a serious propoblem for embedded systems with strict memory constraint.
The proposed framework realizes the componentization of TINET while satisfying the memory constraint, because components statically generate and dynamically connect in TECS.

TINET+TECS utilizes the dynamic connection to swith between CEP and REP components as shown in Fig. \ref{fig:DynamicConnectionUseCase}.
In a server application, CEP is associated with REP in the state waiting for connection request from clients\footnote{ITRON TCP/IP API Specification \cite{url:ITRON_TCP/IP_API_Spec}\\tcp\_acp\_cep(ID cepid, ID repid, T\_IPV4EP *p\_dstaddr, TMO tmout)}.
For example, when processing with HTTP protocol, CEP passively opens with REP of port number 80.

@@

To utilize dynamic connection, the selector should be defined.
The selector connects all the components that can be dynamically connnected, to refer the descripter of them.
Descriptor is an identifier to access the component.
cREP ports are call ports array, which connect the number of tREP cells (Line 9 in Fig. \ref{src:DynamicCDLcode}).
{\it [ref\_desc]} is described to identify the call port refers to descriptors. 
In the case of Fig. \ref{fig:DynamicConnectionUseCase}, the tRepSelector cell connects all tREP cells.

@@

\begin{figure}[t]
\centering
\begin{lstlisting}
eAPI_accept (.., ..) {
    /* Get a descripter of intended REP cell */
    cRepSelector_getRep(&desc, repid);
    /* Set the descripter */
    cREP_set_descriptor(desc);
    /* Call the function of intended REP cell */
    cREP_getEndpoint();
}
\end{lstlisting}
\caption{Accept function (a dynamic connection example)}
\label{src:DynamicCcode}
\end{figure}

@@

\subsection{Evaluation environment}

GR-PEACH is employed as the evaluation board.
We connected the board and the host PC with a LAN cable, and evaluated the data transmission and reception.
The detail specification of the board is shown in TABLE \ref{tab:EvaluationBoardEnvironment}.
We also employ TINET 1.5.4 and the compiler arm-none-eabi--gcc 5.2.

@@

@@

{\bf CiAO/IP:}

@@

In the future, the proposed framework will cooperate with mruby on TECS \cite{par:mrubyonTECS}.
Note that mruby is a scripting language for embedded systems \cite{par:mruby}.
We will support the functionality that TINET functions can be utilized from mruby programs as an extension of mruby-socket \cite{url:mruby-library}.

>>>>

Embedded systems are applied to Internet of Things (IoT), and the high productivity of embedded network software is required.
TINET is a TCP/IP protocol stack for embedded systems.
Although TINET is a compact TCP/IP protocol stack, it consists of many complex source codes.
Therefore, it is difficult to maintain, extend, and analysis the software.
To improve the scalability and configurability, this paper proposes TINET+TECS, a component-based TCP/IP protocol stack for embedded systems: TINET componentized with TOPPERS embedded component system (TECS).
The component-based TINET provides software developers high productivity such as change of network buffer size and adding/removing TCP (or UDP) function.
TINET+TECS utilizes a dynamic connection method of TECS components to satisfy the original TINET specification.
We evaluate the component-based TINET compared with the original TINET.
Experimental results show that the overheads of execution time and memory consumption are low, and that the configurability is improved.

@@

This paper proposes component-based TCP/IP protocol stack, i.e., TINET+TECS, to improve the configurability and scalability of software.
TECS (TOPPERS Embedded Component System) \cite{par:TECS} is utilized to componentize TINET, because TECS is a component system suitable for embedded systems.
TECS supports static configuration, which statically defines component behaviors and interconnections.
Thus, TECS can optimize the overhead of componentization.

In addition, satisfying the original TINET specification, the proposed framework utilizes dynamic connection, a method of TECS, to dynamically switch component bindings.
A general TCP/IP protocol server dynamically processes the requested port, i.e., HTTP (port: 80) and HTTPS (port: 443).
However, embedded systems are restricted dynamic processing due to the strict memory restriction.
TINET supports static generation of CEPs and REPs, which is like sockets.
As TINET+TECS statically generates components and dynamically combines them as well as TINET, TINET+TECS reduces dynamically increasing memory consumption.

In the proposed framework, software applications can be developed by not only TECS method but also existing method.
The software applications can be developed as a TECS component because TINET+TECS is a component-based framework using TECS.
Moreover, TECS supports an adapter to call functions of TECS components from non TECS codes.
The adapter allows the use of existing TCP/IP applications without modification.

@@

This section describes the system model of TINET+TECS, including the basic technologies such as TINET and TECS.
System model of the proposed framework is shown as Fig. \ref{fig:SystemModel}.
TINET+TECS is a component-based TCP/IP protocol stack, so TCP output task (tTCPOutput) and Ethernet input task (tEhternetInput) are implemented as TECS components.
CEP and REP (Section \ref{sec:TINET}), which are also implemented as TECS components, dynamically switch bindings by TECS method.
Moreover, TECS adapter supports the legacy codes for existing applications.

\begin{figure}[t]
    \centering
    \includegraphics[width=7.0cm,clip]{figure/SystemModel.pdf}
    \caption{System model of TINET+TECS}
    \label{fig:SystemModel}
\end{figure}

@@

The components of a protocol stack for TINET+TECS is shown in Fig. \ref{fig:ComponentProtocolStack}.
Note that some small particle components such as a kernel object, dataqueques, and semaphores are omitted to simplify the component diagram.
In TINET+TECS, the components are divided for each protocol, and the functionalities such as input function and output function are defined as each a component.
Therefore, software visibility is improving because of small grain components.

@@

{\bf Application layer:}
An application in TINET+TECS is implemented as a component such as tApplication.
Software with TINET uses ITRON TCP/IP API \cite{url:ITRON_TCP/IP_API_Spec} such as {\it tcp\_snd\_dat} and {\it tcp\_rcv\_dat}.
In TINET+TECS, the application component calls TECS functions such as {\it cTCPAPI\_sendData} and {\it cTCPAPI\_receiveData}.
Moreover, TINET+TECS supporting a TECS adapter (\ref{sec:TECS Adapter}), an existing application with TINET can run on TINET+TECS framework without transporting.
Therefore, software can be developed either with existing methods or as a TECS component.

{\bf Transport layer:}
tTCPCEP (tUDPCEP) is a CEP component for TCP (UDP), and tREP is a REP component.
For example, a server program supporting multiple clients can be developed by preparing the multiple tTCPCEP components.
tTCPInput (tUDPInput) and tTCPOutput (tUDPOutput) are components performing the receiving and sending processing respectively in the transport layer.

@@

To utilize the protocol stack as same as the original TINET, communication object components such as tTCPCEP, tUDPCEP, and tREP are defined as an interface between TINET+TECS and an application.
The communication object component is a component corresponding to a CEP or REP of the original TINET.
Application developers can utilize the TINET+TECS functionalities by generating and combining as many components as necessary.

@@

The original TINET eliminates dynamic memory control to meet the severe memory restriction of embedded systems.
A memory area for sending/receiving data in the protocol stack is allocated and released within a predetermined area.
Memory allocator component performs the elimination of dynamic memory control in TINET+TECS.
The component provides a requested memory area from the memory area statically allocated.

@@

{\it send} is also a specifier for transferring data to a callee from a caller such as {\it in}.
The difference between {\it in} and {\it send} is whether to deallocate the data memory in a caller or callee, shown in Fig. \ref{fig:SendReceive}.
In case of {\it in} specifier, both allocating and deallocating the data memory are performed in the caller.
On the other hand, in case of {\it send}, the caller allocates the data memory and the callee deallocates it.

@@

{\it receive} is also a specifier for a caller receiving data from a callee such as {\it out}.
The difference between {\it out} and {\it receive} is whether to allocate the data memory in a caller or callee, shown in Fig. \ref{fig:SendReceive}.
While, in case of {\it out}, a callee writes data in the memory allocated by a caller, in case of {\it receive}, the callee allocates the data memory.
Deallocating the memory is performed in the caller in both cases.

@@

% \subsubsection{Network timer component}

@@

TECS supports dynamic connection as a new functionality.
Dynamic connection is a method to switch the binding of components at runtime as shown in Fig. \ref{fig:DynamicConnection}.
Note that all components are statically generated in TECS.
TECS can optimize the overhead of componentization because components are statically configured.
Dynamically generating the components causes a lot of memory consumption, which is a serious problem for embedded systems with strict memory constraint.
The proposed framework realizes the componentization of TINET while satisfying the memory constraint, because components statically generate and dynamically connect in TECS.

TINET+TECS utilizes the dynamic connection to switch between CEP and REP components as shown in Fig. \ref{fig:DynamicConnectionUseCase}.
In a server application, CEP is associated with REP in the state waiting for connection request from clients\footnote{ITRON TCP/IP API Specification \cite{url:ITRON_TCP/IP_API_Spec}\\tcp\_acp\_cep(ID cepid, ID repid, T\_IPV4EP *p\_dstaddr, TMO tmout)}.
For example, when processing with HTTP protocol, CEP passively opens with REP of port number 80.

@@

To utilize dynamic connection, the selector should be defined.
The selector connects all the components that can be dynamically connnected, to refer the descriptor of them.
Descriptor is an identifier to access the component \cite{par:optimization}.
cREP ports are call ports array, which connect the number of tREP cells (Line 9 in Fig. \ref{src:DynamicCDLcode}).
{\it [ref\_desc]} is described to identify the call port refers to descriptors. 
In the case of Fig. \ref{fig:DynamicConnectionUseCase}, the tRepSelector cell connects all tREP cells.

@@

\begin{figure}[t]
\centering
\begin{lstlisting}
eAPI_accept (.., ..) {
    /* Get a descriptor of intended REP cell */
    cRepSelector_getRep(&desc, repid);
    /* Set the descriptor */
    cREP_set_descriptor(desc);
    /* Call the function of intended REP cell */
    cREP_getEndpoint();
}
\end{lstlisting}
\caption{Accept function (a dynamic connection example)}
\label{src:DynamicCcode}
\end{figure}

@@

This section describes our experimental evaluation to demonstrate the effectiveness of the proposed framework.

\subsection{Evaluation Environment}

GR-PEACH is employed as the evaluation board.
We connected the board and the host PC with a LAN cable, and evaluated the data transmission and reception.
The detail specification of the board is shown in TABLE \ref{tab:EvaluationBoardEnvironment}.
We also employ TINET 1.5.4 and the compiler arm-none-eabi--gcc 5.2

@@

\subsection{Performance of TINET+TECS}

To demonstrate the low overhead of TINET+TECS, we evaluates the execution time and the memory consumption of TINET+TECS compared with TINET.

Fig. \ref{fig:EvaluationOfExecutionTime} shows the comparison of execution time between TINET and TINET+TECS.
The APIs used for the evaluation are {\it tcp\_snd\_dat} to send TCP data and {\it tcp\_rcv\_dat} to receive TCP data.
For {\it tcp\_snd\_dat}, we measured the executing time from the API calling by the application to the return of the processing result.
In TINET+TECS, the process is performed in the order of tApplication, tTCPCEP, tTCPOutputTaskBody, tIPv4Output, tEthernetOutput, tArp, tEthernetOutputTaskBody, and tIfMbed of Fig. \ref{fig:ComponentProtocolStack}.
For {\it tcp\_rcv\_dat}, we measured the execution time from the data receiving in the LAN driver to the data acquisition in the application.
In TINET+TECS, the process is performed in the order of tIfMbed, tEthernetInputTaskBody, tIPv4Input, tTCPInput, tTCPCEP, and tApplication of Fig. \ref{fig:ComponentProtocolStack}.
The execution time of TINET+TECS is as well as that of TINET; the overhead is about 0000 usec.
{\it send/receive} specifiers enable to access the buffer address without data copies.
Therefore, the overhead of componentization does not effect the execution time.

\begin{figure}[t]
    \centering
    \includegraphics[width=7.0cm,clip]{figure/EvaluationOfExecutionTime.pdf}
    \caption{Execution time of TINET and TINET+TECS}
    \label{fig:EvaluationOfExecutionTime}
\end{figure}

TABLE \ref{tab:EvaluationOfMemoryConsumption} shows the memory consumption of TINET and TINET+TECS.
The memory consumption of TINET+TECS increases about 0000\% compared with TINET.
The processes and data to manage TECS components, such as initialization of cells, descriptors, function tables, and skeleton functions, cause the increase memory consumption.

\begin{table}[t]
    \centering
    \caption{Memory consumption of TINET and TINET+TECS}
    \begin{tabular}{c|r|r}
        \hline\hline
                  &   TINET         &   TINET+TECS   \\ \hline
        RAM       &   10,000 Bytes  &   10,000 Bytes \\
        ROM       &   10,000 Bytes  &   10,000 Bytes \\
        RAM + ROM &   20,000 Bytes  &   20,000 Bytes \\
        \hline
    \end{tabular}
    \label{tab:EvaluationOfMemoryConsumption}
\end{table}

Next, to demonstrate the improved configurability, we measured the code lines for modification as shown in TABLE \ref{tab:EvaluationOfConfigurabilty}.
We can change the composition of the protocol stack with a small workload.
Thus, the proposed framework improves the configurability.

\begin{table}[t]
    \centering
    \caption{Modified code lines of CDL}
    \begin{tabular}{c|r|r|r}
        \hline\hline
                         &   Size          &   Size (- Default)  & CDL  \\ \hline
        Default          &   10,000 Bytes  &   10,000 Bytes      & 0    \\
        I                &   10,000 Bytes  &   10,000 Bytes      & 2    \\
        I + I\hspace{-.1em}I &   20,000 Bytes  &   20,000 Bytes  & 5    \\
        I + I\hspace{-.1em}I + I\hspace{-.1em}I\hspace{-.1em}I   &   20,000 Bytes  &   20,000 Bytes  & 20 \\
        \hline
    \end{tabular}
    \label{tab:EvaluationOfConfigurability}
\end{table}

\subsection{Dynamic Connection}

\begin{figure}[t]
    \centering
    \includegraphics[width=7.0cm,clip]{figure/ComparisonOfDynamicConnection.pdf}
    \caption{Component diagrams of two cases (with/without dynamic connection)}
    \label{fig:ComparisonOfDynamicConnection}
\end{figure}

We evaluated the memory consumption without and with TECS dynamic connection.
As shown in the left of Fig. \ref{fig:ComparisonOfDynamicConnection}, a CEP component should statically connect to all REP components in the case dynamic connection is not used.
As the number of REPs increases, call ports of CEP are required for that.
Thus, it consumes a lot of memory. 
Dynamic connection reduces the memory consumption because only one call port of CEP for REP, which is drawn with red lines in the right of Fig. \ref{fig:ComparisonOfDynamicConnection}, is required per CEP.
Even when the number of REPs increases, a call port of the selector, not the CEP, is joined.

Fig. \ref{fig:EvaluationOfDynamicConnection} shows the memory consumption of two cases, without/with dynamic connection.
The case of dynamic connection consumes the RAM memory more than the other case because the call ports with {\it [dynamic]} is not optimized and allocated in RAM areas as mentioned in Section \ref{sec:DynamicConnection}.
The total memory consumption is fewer in the proposed framework.

\begin{figure}[t]
    \centering
    \includegraphics[width=7.0cm,clip]{figure/EvaluationOfDynamicConnection.pdf}
    \caption{Memory consumption of two cases (with/without dynamic connection)}
    \label{fig:EvaluationOfDynamicConnection}
\end{figure}

\subsection{Overhead of Adapter}

TINET+TECS supports existing applications because of the TECS adapter.
To analyze the overhead of TECS adapter, we measured execution time of API with the adapter such as {\it tcp\_snd\_dat/eAPI\_sendData}.
As shown in Fig. \ref{fig:EvaluationOfAdapter}, the overhead of TECS adapter is very small because the adapter only passes the parameters to TECS components from C codes.
Therefore, TECS adapter does not affect the system.

\begin{figure}[t]
    \centering
    \includegraphics[width=7.0cm,clip]{figure/EvaluationOfAdapter.pdf}
    \caption{Execution time of two cases (without/with TECS adapter)}
    \label{fig:EvaluationOfAdapter}
\end{figure}

@@

{\bf CiAO/IP:}
\\
\\
Should be written
\\
\\

@@

This paper proposes TINET+TECS, a component-based TCP/IP protocol stack for embedded systems.
TINET+TECS is componentized TINET, which is a compact TCP/IP protocol stack, using TECS.
TINET consists of many macros and complicated codes and the software productivity are low.
The proposed framework improves the configurability while suppressing the overhead of componentization.
The proposed framework also improves the scalability because the component-based framework simplifies to add/remove and change protocols such as TCP/UDP, IPv4/IPv6, and Ethernet/PPP.

In addition, this paper presents dynamic connection, a new method of TECS.
The method enables dynamic processing while reducing memory consumption.
To satisfy TINET specification that TINET support static generation of CEPs and REPs, TINET+TECS utilizes dynamic connection.
TECS adapter supports legacy codes, existing TCP/IP applications can run without modification in the proposed framework.

In the future, the proposed framework will cooperate with mruby on TECS \cite{par:mrubyonTECS} to manage IoT devices.
Note that mruby is a scripting language for embedded systems \cite{par:mruby}.
We will support the functionality that TINET functions can be utilized from mruby programs as an extension of mruby-socket \cite{url:mruby-library}.
